---
title: "Getting started with YieldLag"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with YieldLag}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(yieldlag)
```

YieldLag estimates regional crop-yield associations with seasonal weather.
The workflow separates data construction, tuning, evaluation, fitting, and prediction.

## Seasonal nonlinear example

```{r}
data <- synthetic_stress_example()
model <- fit_crop_model(data, "stress_lag", parameters = list(
  lambda = 1, smooth_ratio = 0, threshold = 0.75,
  temperature_variable = "TMP", precipitation_variable = "PCP"),
  robust = FALSE)
future <- subset(data$panel, Year == 2031)
head(predict(model, future))
```

This fictional dataset has a declared asymmetric June response and hot-dry
interaction. It illustrates the model, not observed-crop skill. Direct fitting
uses fixed parameters and does not calibrate a prediction interval.

## Interpretation and support

```{r}
explain_crop_prediction(model, future)
prediction_support(model, future)
```

Components sum to the fitted prediction relative to training feature means.
They express statistical associations, not causal losses. Support diagnostics
count weather terms outside training ranges and flag unseen regions.

## Chronological comparison

```{r eval=FALSE}
config <- crop_model_config(
  methods = c("trend", "panel_ridge", "stress_lag", "ensemble"),
  weather_variables = c("TMP", "PCP"),
  stress_temperature = "TMP", stress_precipitation = "PCP", robust = FALSE
)
comparison <- compare_crop_models(data, config, 2030, 2031)
summary(comparison)
selected <- get_crop_model(comparison)
predict(selected, future, interval = "prediction")
write_crop_results(comparison, "results/example")
```

Automatic selection uses tuning RMSE. Exports include candidate scores, predictions,
metrics, nonlinear basis coefficients, stress components, support diagnostics,
fitted objects, configuration, and checksums.

## Observed data

```{r}
wheat <- read_crop_data(poland_example_path(), "wheat", harvest_month = 7)
wheat
```

Use `prepare_crop_data()` for other panel schemas with explicit column mappings.
The Poland example contains monthly study tables and fixed areas. Complete
harvest-year weather defines a retrospective benchmark; source terms and
metadata limitations accompany the data.
