---
title: "Getting started with cropyieldmodel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with cropyieldmodel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(cropyieldmodel)
```

`cropyieldmodel` separates data preparation, declared fitting, chronological
comparison, model selection, and artifact writing. That separation is
intentional: an in-sample fit is not presented as validation.

## Read the reproducible example

```{r}
wheat <- read_crop_data(
  poland_example_path(), crop = "wheat", harvest_month = 7
)
wheat
```

The reader constructs 84 unselected monthly terms: seven declared weather
variables by twelve months. To use another layout, call `prepare_crop_data()`
and map the panel columns explicitly.

## Fit a declared model

```{r}
trend <- fit_crop_model(wheat, method = "trend")
trend
head(predict(trend, wheat$panel))
```

Direct fitting uses fixed settings and intentionally has no calibrated
prediction interval.

## Compare methods without temporal leakage

The complete workflow below selects hyperparameters in 2006--2011, evaluates
rolling forecasts in 2012--2018, fits final models through 2018, and predicts
2019. It is not evaluated while building the vignette because the full grid is
a substantive model run.

```{r eval=FALSE}
comparison <- compare_crop_models(
  wheat,
  config = crop_model_config(),
  last_yield_year = 2018,
  predict_through = 2019
)

comparison
summary(comparison)
plot(comparison, level = "Province-year")

selected <- get_crop_model(comparison)
predict(
  selected,
  wheat$panel[wheat$panel$Year == 2019, ],
  interval = "prediction"
)
```

`get_crop_model(comparison)` returns the method selected by tuning RMSE only.
It does not select the best-looking method from later evaluation results.

## Export every decision and prediction

```{r eval=FALSE}
write_crop_results(comparison, "results/wheat", overwrite = FALSE)
```

The result bundle contains accepted and rejected hyperparameter candidates,
tuning and evaluation predictions, metrics, final models, configuration in
text and R formats, input and output checksums, R session information, and a
plain-language report.
